A memory frontier for complex synapses

### 2013-05-31 NIPS/SynapseNIPS.tex
fixed typo, back to svg

\begin{figure}[tbp]
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=0.25\linewidth]{cascade-generated.pdf}}\label{fig:cascade}
  \item\aligntop{\includegraphics[width=0.25\linewidth]{multistate-generated.pdf}}\label{fig:multistate}
  \item\aligntop{\includegraphics[width=0.35\linewidth]{cascms.eps}}\label{fig:cascms}
 \end{myenuma}
 \end{center}
  \caption{Models of complex synapses.
  (\ref{fig:cascade}) The cascade model of \cite{Fusi2005cascade}, showing transitions between states of high/low synaptic weight (red/blue circles) due to potentiation/depression (solid red/dashed blue arrows).
  (\ref{fig:multistate}) The serial model of \cite{Leibold2008serial}.
  (\ref{fig:cascms}) The memory curves of these two models, showing the decay of the signal-to-noise ratio (to be defined in \autoref{sec:setup}) as subsequent memories are stored.
  }\label{fig:models}
\end{figure}

@@

This two state model has the equilibrium distribution $\eq=(f\dep,f\pot)$ and its flux is given by $\F_{-+} = rf\pot f\dep$.
This is maximized when $f\pot=f\dep=\half$, leading to the upper bound:
%
\begin{equation}\label{eq:max_init}
  \snr(0) \leq \sqrt{N}.
\end{equation}
%
We note that while this model has high initial SNR, it also has very fast memory decay -- with a timescale $\tau\sim\frac{1}{r}$.
As the synapse is very plastic, the initial memory is encoded very easily, but the subsequent memories also overwrite it rapidly.
This is one example of the tradeoff between optimizing memory at early versus late times.
\begin{figure}[tbp]
 \begin{center}
 \begin{myenuma}
  \item\hp\aligntop{\includegraphics[width=0.15\linewidth]{max_init_pot-generated.pdf}}\label{fig:max_init_pot}\hp
  \item\aligntop{\includegraphics[width=0.15\linewidth]{max_init_dep-generated.pdf}}\label{fig:max_init_dep}\hp
  \item\hp\aligntop{\includegraphics[width=0.05\linewidth]{binary_det-generated.pdf}}\label{fig:binary_det}
  \end{myenuma}
 \end{center}
  \caption{Synaptic models that maximize initial SNR.
  (\ref{fig:max_init_pot}) For potentiation, all transitions starting from a weak state lead to a strong state, and the probabilities for all transitions leaving a given weak state sum to 1.
  (\ref{fig:max_init_pot}) Depression is similar to potentiation, but with strong and weak interchanged.
  (\ref{fig:binary_det}) The equivalent two state model, with transition probabilities under potentiation and depression equal to one.
  }\label{fig:max_init}
\end{figure}

@@

Our goal of understanding the relationship between structure and function in the space of all possible synaptic models is complicated by the fact that this space contains many different possible network topologies, encoded in the nonzero matrix elements of $\M\potdep$.  To systematically analyze this entire space, we develop an important organizing principle using the theory of first passage times in the stochastic process of forgetting, described by $\frg$.
The mean first passage time matrix, $\fptb_{ij}$, is defined as the average time it takes to reach state $j$ for the first time, starting from state $j$.
The diagonal elements are defined to be zero.

@@

\begin{figure}[tbp]
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=0.1\linewidth]{triangular_right-generated.pdf}}\label{fig:tri_right}\hp
  \item\aligntop{\includegraphics[width=0.1\linewidth]{triangular_left-generated.pdf}}\label{fig:tri_left}\hp
  \item\aligntop{\includegraphics[width=0.2\linewidth]{shortcut-generated.pdf}}\label{fig:shortcut}
 \end{myenuma}
 \end{center}
  \caption{Perturbations that increase the area.
  (\ref{fig:tri_right}) Perturbations that increase elements of $\M\pot$ above the diagonal and decrease the corresponding elements of $\M\dep$. It can no longer be used when $\M\dep$ is lower triangular.
  (\ref{fig:tri_left}) Perturbations that decrease elements of $\M\pot$ below the diagonal and increase the corresponding elements of $\M\dep$. It can no longer be used when $\M\pot$ is upper triangular.
  (\ref{fig:shortcut}) Perturbation that decreases ``shortcut'' transitions and increases the bypassed ``direct'' transitions. It can no longer be used when there are only nearest-neighbor ``direct'' transitions.
  }\label{fig:perts}
\end{figure}

@@

\begin{figure}[tbp]
 \begin{center}
  \includegraphics[width=0.35\linewidth]{multistate_sticky-generated.pdf}
 \end{center}
  \caption{Model that maximizes the area.
  Unlabeled transitions have probability 1.
  Labeled transitions have probability $\varepsilon\to0$.
  }\label{fig:max_area}
\end{figure}

@@

\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=0.38\linewidth]{env.eps}}\label{fig:envplot}\hp
  \item\aligntop{\includegraphics[width=0.3\linewidth]{diffjump-generated.pdf}}\label{fig:diffjump}
 \end{myenuma}
 \end{center}
  \caption{The memory curve envelope for $N=100$, $M=12$.
  (\ref{fig:envplot}) An upper bound on the SNR at any time is shown in green.
  The red dashed curve shows the result of numerical optimization of synaptic models with random initialization.
  The solid red curve shows the highest SNR we have found with hand designed models.
  At early times these models are of the form shown in (\ref{fig:diffjump}) with different numbers of states, and all transition probabilities equal to $1$.
  At late times they are of the form shown in \autoref{fig:max_area} with different values of $\varepsilon$.
  }\label{fig:env}
\end{figure}

@@

%\subsubsection*{Acknowledgements}
%We thank Sloan, Genenetech, Burroughs-Wellcome, and Swartz foundations for support.

>>>>

\begin{figure}[tbp]
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=0.25\linewidth]{cascade.svg}}\label{fig:cascade}
  \item\aligntop{\includegraphics[width=0.25\linewidth]{multistate.svg}}\label{fig:multistate}
  \item\aligntop{\includegraphics[width=0.35\linewidth]{cascms.eps}}\label{fig:cascms}
 \end{myenuma}
 \end{center}
  \caption{Models of complex synapses.
  (\ref{fig:cascade}) The cascade model of \cite{Fusi2005cascade}, showing transitions between states of high/low synaptic weight (red/blue circles) due to potentiation/depression (solid red/dashed blue arrows).
  (\ref{fig:multistate}) The serial model of \cite{Leibold2008serial}.
  (\ref{fig:cascms}) The memory curves of these two models, showing the decay of the signal-to-noise ratio (to be defined in \autoref{sec:setup}) as subsequent memories are stored.
  }\label{fig:models}
\end{figure}

@@

This two state model has the equilibrium distribution $\eq=(f\dep,f\pot)$ and its flux is given by $\F_{-+} = rf\pot f\dep$.
This is maximized when $f\pot=f\dep=\half$, leading to the upper bound:
%
\begin{equation}\label{eq:max_init}
  \snr(0) \leq \sqrt{N}.
\end{equation}
%
We note that while this model has high initial SNR, it also has very fast memory decay -- with a timescale $\tau\sim\frac{1}{r}$.
As the synapse is very plastic, the initial memory is encoded very easily, but the subsequent memories also overwrite it rapidly.
This is one example of the tradeoff between optimizing memory at early versus late times.
\begin{figure}[tbp]
 \begin{center}
 \begin{myenuma}
  \item\hp\aligntop{\includegraphics[width=0.15\linewidth]{max_init_pot.svg}}\label{fig:max_init_pot}\hp
  \item\aligntop{\includegraphics[width=0.15\linewidth]{max_init_dep.svg}}\label{fig:max_init_dep}\hp
  \item\hp\aligntop{\includegraphics[width=0.05\linewidth]{binary_det.svg}}\label{fig:binary_det}
  \end{myenuma}
 \end{center}
  \caption{Synaptic models that maximize initial SNR.
  (\ref{fig:max_init_pot}) For potentiation, all transitions starting from a weak state lead to a strong state, and the probabilities for all transitions leaving a given weak state sum to 1.
  (\ref{fig:max_init_pot}) Depression is similar to potentiation, but with strong and weak interchanged.
  (\ref{fig:binary_det}) The equivalent two state model, with transition probabilities under potentiation and depression equal to one.
  }\label{fig:max_init}
\end{figure}

@@

Our goal of understanding the relationship between structure and function in the space of all possible synaptic models is complicated by the fact that this space contains many different possible network topologies, encoded in the nonzero matrix elements of $\M\potdep$.  To systematically analyze this entire space, we develop an important organizing principle using the theory of first passage times in the stochastic process of forgetting, described by $\frg$.
The mean first passage time matrix, $\fptb_{ij}$, is defined as the average time it takes to reach state $j$ for the first time, starting from state $i$.
The diagonal elements are defined to be zero.

@@

\begin{figure}[tbp]
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=0.1\linewidth]{triangular_right.svg}}\label{fig:tri_right}\hp
  \item\aligntop{\includegraphics[width=0.1\linewidth]{triangular_left.svg}}\label{fig:tri_left}\hp
  \item\aligntop{\includegraphics[width=0.2\linewidth]{shortcut.svg}}\label{fig:shortcut}
 \end{myenuma}
 \end{center}
  \caption{Perturbations that increase the area.
  (\ref{fig:tri_right}) Perturbations that increase elements of $\M\pot$ above the diagonal and decrease the corresponding elements of $\M\dep$. It can no longer be used when $\M\dep$ is lower triangular.
  (\ref{fig:tri_left}) Perturbations that decrease elements of $\M\pot$ below the diagonal and increase the corresponding elements of $\M\dep$. It can no longer be used when $\M\pot$ is upper triangular.
  (\ref{fig:shortcut}) Perturbation that decreases ``shortcut'' transitions and increases the bypassed ``direct'' transitions. It can no longer be used when there are only nearest-neighbor ``direct'' transitions.
  }\label{fig:perts}
\end{figure}

@@

\begin{figure}[tbp]
 \begin{center}
  \includegraphics[width=0.35\linewidth]{multistate_sticky.svg}
 \end{center}
  \caption{Model that maximizes the area.
  Unlabeled transitions have probability 1.
  Labeled transitions have probability $\varepsilon\to0$.
  }\label{fig:max_area}
\end{figure}

@@

\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=0.38\linewidth]{env.eps}}\label{fig:envplot}\hp
  \item\aligntop{\includegraphics[width=0.3\linewidth]{diffjump.svg}}\label{fig:diffjump}
 \end{myenuma}
 \end{center}
  \caption{The memory curve envelope for $N=100$, $M=12$.
  (\ref{fig:envplot}) An upper bound on the SNR at any time is shown in green.
  The red dashed curve shows the result of numerical optimization of synaptic models with random initialization.
  The solid red curve shows the highest SNR we have found with hand designed models.
  At early times these models are of the form shown in (\ref{fig:diffjump}) with different numbers of states, and all transition probabilities equal to $1$.
  At late times they are of the form shown in \autoref{fig:max_area} with different values of $\varepsilon$.
  }\label{fig:env}
\end{figure}

@@

%\subsubsection*{Acknowledgements}
%We thank Sloan, Genenetech, Burroughs-Wellcome, and Swartz foundations for support.
%We thank larry Abbott, Marcus Benna, Stefano Fusi, Jascha Sohl-Dickstein and David Sussillo for useful discussions.
